Statistical Modeling of Galaxy Catalogues with Normalizing Flows

### 2022-08-17 src/tex/ms.tex
Updated pz ensemble and point estimate plot.

For example, RQ-RSCs (and the RQ-NSCs on which they are based) are defined on the domain [-B, B], and thus will not transform samples outside this range.
It is therefore useful to define a \emph{Shift Bounds} bijection, which shifts the original range of each dimension to match the domain of the splines.
Note that this shift must be set at training time, with the assumption that future test data will lie within the same bounds\footnote{This bijection may be poorly suited to data with large outliers. In that case, more care should be taken in designing a bijection that maps the data onto the range of the splines.}.
You can choose a range wider than that covered by the training set if you wish to allow the flow to sample outside the range of the training set.

\note{Discuss the Inverse Softplus Bijection here! If I use it below. Not sure if I will...}

@@

\note{
Discuss latent distributions here.
Note that most applications use a Normal Distribution, but since the splines have a compact domain, we need to choose a latent distribution with compact support, or else we might generate strange outliers when sampling from the flow.
Some text for this might already be in the creation module draft.
}

@@

See Figure \ref{fig:sphere} for an example of circular splines used to model the population density of the Earth \note{cite the data set here}.
The angles are shifted so that $\phi=\pm \pi$ corresponds to the Prime Meridian.
You can see that the circular splines have no problem with smoothly modeling the data around this discontinuity in $\phi$.

@@

\note{Try to get the discrete example working again! Add a figure here.}

@@

PZFlow is a Python package for building normalizing flows, with a focus on easy high-performance modeling of high-dimensional tabular data.
Data is handled in Pandas DataFrames \citep{pandas}, while the normalizing flows are implemented in Jax \citep{jax}, which allows for efficient, parallelizable, GPU-enabled calculations for very large data sets.
The code is easily installable for the Python Package Index\footnote{\url{https://pypi.org/project/pzflow/}} (PyPI) and is hosted on Github,\footnote{\url{https://github.com/jfcrenshaw/pzflow}}.
The documentation\footnote{\note{URL}} includes tutorial notebooks covering all of the features demonstrated in this paper.

@@

In addition to the examples in this paper, PZFlow has already being used in various other projects:
\begin{itemize}
    \item \needscite used PZFlow to build a metric for observing strategy optimization based on information theory;
    \item \needscite used PZFlow to forward model galaxy data with true redshift posteriors in order to evaluate the impact of survey incompleteness and spec-z errors on photo-z estimation;
    \item \needscite used PZFlow to smooth high-redshift artifacts in simulations of supernova host galaxy simulations.
\end{itemize}

@@

In Section \ref{sec:fwd-model} we construct a normalizing flow to model the galaxy redshifts and photometry.
In Section \note{LABEL}, we calculate posteriors?
In Section \ref{sec:fwd-model-conditional} we build a conditional flow to add additional properties to the catalog.

@@

To create a generative model of galaxy redshifts and photometry, we use the true redshifts and $ugrizy$ magnitudes from the DESC CosmoDC2 simulation (\needscite).
We selected all galaxies from CosmoDC2 with at most one band beyond the LSST 10-year 5-sigma point source depths.
Of these, we selected $10^6$ galaxies and split them intro training and test sets consisting of 80\% and 20\% of the galaxies, respectively.

For the latent distribution we use a 7 dimensional Uniform distribution over the range $[-5, 5]$.
To map the data onto the latent distribution, we use the following bijection:
\begin{align}
    f = \text{RQ-RSC} \circ \text{Shift Bounds} \circ \text{Color Transform}.
\end{align}
\note{I might add an InverseSoftPlus}
We will explain each layer of the bijection in the order they are applied to the input data.

@@

We trained the normalizing flow for 150 epochs using the Adam optimizer \citep{adam} (with a learning rate of $10^{-3}$, that decreased by a factor of 10 every 50 epochs) to minimize the negative log-likelihood,
\begin{align}
    \mathcal{L} = - \, \mathbb{E}[ \, \log p(x) \, ],
\end{align}
where the expectation is performed over galaxies in the training set and $p(x)$ is defined in Equation \ref{eq:px}.
Training took 7 minutes on a Tesla P100 12GB GPU.

We assess the results by drawing $10^4$ galaxies from the trained flow, and plotting their distribution against  $10^4$ galaxies from the test set (Figure \ref{fig:main-corner}).
We see that the normalizing flow has done an excellent job of reproducing the distribution of DESC DC2, without any unusual artifacts or outliers.
In addition, Figure \ref{fig:smooth-color-dist} compares the distribution of galaxy $r-i$ vs redshift.
The CosmoDC2 simulation is known to exhibit discrete tracks in this space at high redshift, due to the discrete number of SED templates used during simulation.
These tracks are visible in the left panel.
The right panel shows that PZFlow smooths over this discreteness, resulting in a color distribution that is smooth up to high redshifts.

@@

We can also marginalize over any missing bands, $n$:
\begin{align}
    p(z|\hat{m}) = \frac{1}{p(\hat{m})} \int p(z, n, \hat{m}) dn,
\end{align}
which can be calculated by evaluating $p(z, n, \hat{m})$ on a grid of $z$ and $n$, summing over $n$ to yield $p(z, \hat{m})$, and normalizing with respect to redshift to yield $p(z|\hat{m})$.
You can once again average over $\hat{m}$ samples to convolve the photometric errors.
You may wish to marginalize over all values of $n$ if the galaxy was not observed in that band.
This might occur, for example, if you include photometry from Euclid, which will not have complete coverage of the LSST catalog \needscite.
You may also wish to marginalize over all values beyond a limiting magnitude if the galaxy was observed, but not detected in that band.
This might occur, for example, in the low wavelength bands of Lyman dropout galaxies observed by LSST.

@@

\subsection{Additional properties with conditional flows}
\label{sec:fwd-model-condtional}

@@

The conditional flow is trained for \note{N} epochs using the same loss function and optimizer as before.
\note{is the LR the same?}
We sample a size and ellipticity for each galaxy in the PZFlow catalog created in the previous section (using the true magnitudes), and plot the distribution of these features against the distribution in the test set (Figure \ref{fig:conditional-corner}).
Once again, we see that the normalizing flow does a good job of emulating the CosmoDC2 distribution.

The final simulated catalog consists of $10^4$ galaxies, each with a redshift, $ugrizy$ magnitudes including photometric errors, a true photo-z posterior consistent with the photometric errors, a size, and an ellipticity.
This small catalog was generated for visualization purposes, but the normalizing flows can be used to generate catalogs of arbitrarly large size (see Appendix \ref{app:benchmarking} on benchmarking PZFlow).

@@

In addition to forward modeling, normalizing flows are powerful and flexible models for density estimation.
This makes them useful tools for estimating posterior distributions for galaxy properties, conditioned on observed properties of the galaxy.
A common example of this in cosmology is photometric redshift estimation, in which you try to estimate the redshift of a galaxy using its magnitude in several photometric bands.
In this section, we will demonstrate this application to show how normalizing flows can be used for density estimation.

@@

A full marginalization over the model parameters would be too computationally expensive, so instead we opt for an ensemble of normalizing flows.
In other words, we will train multiple normalizing flows under identical conditions, just with different random initializations of the model parameters.
This will allow the optimization algorithm to explore different basins of attraction in the parameter space, allowing us to marginalize over different potential models.
In the machine learning literature, this is known as a Deep Ensemble \needscite, a popular method for approximate bayesian marginalization \needscite.

@@

For the training set, we use 100,000 galaxies from the catalog created in Section \ref{sec:galaxy-catalog}.
Each galaxy in the training set has a true redshift, and observed magnitudes in the $ugrizy$ bands, with corresponding photometric errors.
To account for the photometric error, we resample the training set from the error distributions on each epoch.
In other words, each epoch, we resample
\begin{align}
    m \sim p(\hat{m}, \sigma_m),
\end{align}
where $\hat{m}$ are the observed magnitudes with photometric errors $\sigma_m$, and $p(\hat{m}, \sigma_m)$ is a Gaussian in flux space.
This allows our ensemble of flows to approximate the distribution $p(z, m)$.

@@

\begin{figure*}[t!]
    \script{photo-z/plot_ensemble_posteriors.py}
    \begin{centering}
        \includegraphics{figures/ensemble_posteriors.pdf}
        \caption{
            \note{Write this caption}.
        }
        \label{fig:ensemble-posteriors}
    \end{centering}
\end{figure*}

Each flow estimates the posterior by marginalizing over the error distribution.
We then average the posteriors from each flow in the ensemble.
An example of this can be seen in Figure \ref{fig:ensemble-posteriors}.
Each flow produces a PDF which may contain slightly different features in each case.
By averaging over the individual posteriors, we select for features that are common between models, while smoothing over features that are present in only a single model.
Compare these to the true posteriors.

We can calculate posteriors for galaxies with missing magnitudes by marginalizing over the missing magnitudes.
This might be useful when not every galaxy is observed in every band, in which case you can marginalize over the bands that were not observed, or in cases when a galaxy was observed but not detected in the band.
You can then marginalize over values for the band that are beyond the limiting magnitude.
This can also help you understand what information is included in each band and how the inclusion of that band improves photo-z estimation.

For example, we randomly select \note{N} galaxies and estimate redshift posteriors while marginalizing over the $u$ band.
These posteriors can be seen in Figure \note{N}.
Mention how the loss of information degrades the posteriors.

At some point in this section, note that we estimate posteriors for different galaxies than we trained on.

@@

\begin{figure}[t!]
    \script{photo-z/plot_pz_point_estimates.py}
    \begin{centering}
        \includegraphics{figures/pz_point_estimates.pdf}
        \caption{
            \note{Write this caption}.
        }
        \label{fig:point-estimates}
    \end{centering}
\end{figure}

The most common point of comparison is photo-z point estimates.
We make the common choice of selecting the mode of the posteriors.
\note{something something.}

\begin{figure}[t!]
    \script{photo-z/plot_pit_histogram.py}
    \begin{centering}
        \includegraphics{figures/pit_histogram.pdf}
        \caption{
            \note{Write this caption}.
        }
        \label{fig:pit-histogram}
    \end{centering}
\end{figure}

PIT metric.
\note{something something.}
Discuss the mode covering behavior of normalizing flows and how this relates to the PIT diagram.

\begin{figure*}[t!]
    \script{photo-z/plot_binned_metrics.py}
    \begin{centering}
        \includegraphics{figures/binned_metrics.pdf}
        \caption{
            \note{Write this caption}.
        }
        \label{fig:binned-metrics}
    \end{centering}
\end{figure*}

Binned point metrics.
Figure \ref{fig:binned-metrics}
Defined in the science book.
We plot against the requirements from the LSST SRD just to provide a reference for the scale.
Seems to do well up to z~1.5.
We note that for cosmology, it is okay for the bias to exceed the required limits, as long as the bias can be well determined via some calibration process, e.g. clustering redshifts.

The previous metrics concerned photo-z point estimates, which are insufficient for modern science.
The DC1 paper demonstrated that these metrics are not sufficient to validate the full posteriors produced by photo-z codes.
The purpose of this paper is to produce catalogs with true posteriors, which we have done, so that you can evaluate metrics of the full posterior.
We will not undertake a full analysis of the posteriors here, but will compute a few metrics to point in that direction.

Calculate the KLD and maybe something else?

@@

\begin{figure*}[t!]
    \script{forward_model/plot_galaxy_losses.py}
    \begin{centering}
        \includegraphics{figures/galaxy_flow_losses.pdf}
        \caption{
            Training losses for the galaxy flows.
        }
        \label{fig:galaxy-flow-losses}
    \end{centering}
\end{figure*}

@@

- Explain how I adapted the LSST Error Model to work for low-SNR galaxies

\section{Benchmarking PZFlow}
\label{app:benchmarking}

\bibliography{references.bib}

>>>>

For example, RQ-RSCs (and the RQ-NSCs on which they are based) are defined on the domain [-B, B], and thus will not transform samples outside this range.
It is therefore useful to define a \emph{Shift Bounds} bijection, which shifts the original range of each dimension to match the domain of the splines.
Note that this shift must be set at training time, with the assumption that future test data will lie within the same bounds\footnote{This bijection may be poorly suited to data with large outliers. In that case, more care should be taken in designing a bijection that maps the data onto the range of the splines.}.
You can choose a range wider than that covered by the training set if you wish to allow the flow to sample outside the range of the training set

@@

In principle, with a sufficiently expressive bijection, the choice of latent distribution does not matter as long as it is a distribution in which you can easily sample and calculate densities.
However, in practice, bijections are limited in expressiveness, i.e. they cannot necessarily transform any arbitrary data distribution into any arbitrary latent distribution.

For example, the splines of RQ-RSCs only transform samples in the range [-B, B].
Sampling from a latent distribution with support outside this range will therefore result in strange outliers and incorrect boundary conditions.
One can apply a transformation to the latent samples before they are fed into the RQ-RSC to ensure that they lie within the support of the splines, but it is simpler to use a compact latent distribution whose support matches that of the spline layers.
A simple choice would be the uniform distribution over [-B, B].

Additionally, as no bijector is perfect, the structure of the latent distribution will not be completely erased in the translation from latent to data distribution.
Thus, the latent distribution can be viewed as a prior or inductive bias on samples from the data distribution \citep{jaini2020}.
It is therefore advantageous to select a latent distribution whose features match some of the structure in the data.

A latent distribution that can achieve both desiderata is the Beta distribution, i.e. $u \sim \mathrm{Beta}(\alpha, \beta)$, where $\alpha, \beta > 0$ are learnable parameters\footnote{In practice, it is easier to learn $\log\alpha$ and $\log\beta$ to ensure that $\alpha, \beta > 0$.}.
This distribution is compact, and by varying $\alpha$ and $\beta$ this distribution can take on a wide variety of shapes with different means, skews, and kurtoses, allowing the inductive bias of the prior to adapt to structure in the data during training.
However, as RQ-RSCs are defined on the domain $[-B, B]$, it is more convenient to use a modified Beta distribution, which we name the \emph{Centered Beta distribution}:
\begin{align}
    \text{CentBeta}(u | \alpha, \beta, B) = 2B\left(\text{Beta}(u|\alpha, \beta) - \frac{1}{2}\right).
\end{align}

In general, as long as sampling and density evaluation are tractable, one can use any parameterization of the latent distribution that matches some desired structure in the data and learn the distribution parameters during training.
We give this generalization the name \emph{latent-adaptive flows} (LAFs; inspired by the Tail Adaptive Flows of \citealt{jaini2020}).
Our experiments indicate that learnable latent distributions can improve training loss, but require more care in training.

Note that while we discussed univariate distributions above, these considerations generalize easily to multiple dimensions.
Each of these distributions have multivariate generalizations that can be used when modeling higher-dimensional data.
The full multivariate latent distribution can also be assembled by taking the product of multiple univariate distributions\footnote{Note that while the latent variables will be independent, the data variables will still have correlations imprinted by the bijections.}.
This may even be desired if different dimensions of the data have different structure that you wish to encode in the latent distribution.

@@

@@

@@

PZFlow is a Python package for building normalizing flows, with a focus on easy high-performance modeling of high-dimensional tabular data.
Data is handled in Pandas DataFrames \citep{pandas}, while the normalizing flows are implemented in Jax \citep{jax}, which allows for efficient, parallelizable, GPU-enabled calculations for very large data sets.
The code is easily installable for the Python Package Index\footnote{\url{https://pypi.org/project/pzflow/}} (PyPI) and is hosted on Github,\footnote{\url{https://github.com/jfcrenshaw/pzflow}}.
The documentation\footnote{\url{https://jfcrenshaw.github.io/pzflow/}} includes tutorial notebooks demonstrating the features mentioned in this paper on different example problems.

@@

In addition to the examples in this paper, PZFlow has already being used in various other projects:
\begin{itemize}
    \item \citet{malz2021} used PZFlow to build a metric for observing strategy optimization based on information theory;
    \item \citet{stylianou2022} used PZFlow to forward model galaxy data with true redshift posteriors in order to evaluate the impact of survey incompleteness and spec-z errors on photo-z estimation;
    \item \citet{lokken2022} used PZFlow to smooth high-redshift artifacts in simulations of host galaxies for supernovae and other transients.
\end{itemize}

@@

In Section \ref{sec:fwd-model} we construct a normalizing flow to model the galaxy redshifts and photometry and generate a new simulated catalog.
In Section \ref{sec:true-posteriors}, we calculate true redshift posteriors for the new catalog.
In Section \ref{sec:fwd-model-conditional} we build a conditional flow to add additional properties to the catalog.

@@

To create a generative model of galaxy redshifts and photometry, we use the true redshifts and $ugrizy$ magnitudes from the DESC CosmoDC2 simulation \citep{dc2, cosmodc2}.
We selected all galaxies from CosmoDC2 with at most one band beyond the LSST 10-year 5-sigma point source depths.
Of these, we selected $10^6$ galaxies and split them intro training and test sets consisting of 80\% and 20\% of the galaxies, respectively.

For the latent distribution we use a 7 dimensional Uniform distribution over the range $[-5, 5]$.
To map the data onto the latent distribution, we use the following bijection:
\begin{align}
    f = \text{RQ-RSC} \circ \text{Shift Bounds} \circ \text{Color Transform}.
\end{align}
We will explain each layer of the bijection in the order they are applied to the input data.

@@

After training the flow (see Appendix \ref{app:training-details}), we assess the results by drawing $10^4$ galaxies from the trained flow, and plotting their distribution against  $10^4$ galaxies from the test set (Figure \ref{fig:main-corner}).
We see that the normalizing flow has done an excellent job of reproducing the distribution of DESC DC2, without any unusual artifacts or outliers.
In addition, Figure \ref{fig:smooth-color-dist} compares the distribution of galaxy $r-i$ vs redshift.
The CosmoDC2 simulation is known to exhibit discrete tracks in this space at high redshift, due to the discrete number of SED templates used during simulation.
These tracks are visible in the left panel.
The right panel shows that PZFlow smooths over this discreteness, resulting in a color distribution that is smooth up to high redshifts.

@@

We can also marginalize over any missing bands, $n$:
\begin{align}
    p(z|\hat{m}) = \frac{1}{p(\hat{m})} \int p(z, n, \hat{m}) dn,
\end{align}
which can be calculated by evaluating $p(z, n, \hat{m})$ on a grid of $z$ and $n$, summing over $n$ to yield $p(z, \hat{m})$, and normalizing with respect to redshift to yield $p(z|\hat{m})$.
You can once again average over $\hat{m}$ samples to convolve the photometric errors.
You may wish to marginalize over all values of $n$ if the galaxy was not observed in that band.
This might occur, for example, if you include photometry from Euclid, which will not have complete coverage of the LSST catalog \citep{euclid}.
You may also wish to marginalize over all values beyond a limiting magnitude if the galaxy was observed, but not detected in that band.
This might occur, for example, in the low wavelength bands of Lyman dropout galaxies observed by LSST.

@@

\subsection{Additional properties with conditional flows}
\label{sec:fwd-model-conditional}

@@

After training the flow (see Appendix \ref{app:training-details}), we sample a size and ellipticity for each galaxy in the PZFlow catalog created in the previous section (using the true magnitudes), and plot the distribution of these features against the distribution in the test set (Figure \ref{fig:conditional-corner}).
Once again, we see that the normalizing flow does a good job of emulating the CosmoDC2 distribution.

The final simulated catalog consists of $10^4$ galaxies, each with a redshift, $ugrizy$ magnitudes including photometric errors, a true photo-z posterior consistent with the photometric errors, a size, and an ellipticity.
This small catalog was generated for visualization purposes, but the normalizing flows can be used to generate catalogs of arbitrarily large size.

@@

In addition to forward modeling, normalizing flows are powerful and flexible models for density estimation.
This makes them useful tools for estimating posterior distributions for galaxy properties, conditioned on observed features of the galaxy.
A common example of this in cosmology is photometric redshift estimation, in which you try to estimate the redshift of a galaxy using its magnitude in several photometric bands.
In this section, we will demonstrate this application to show how normalizing flows can be used for density estimation.

@@

A full marginalization over the model parameters would be too computationally expensive, so instead we opt for an ensemble of normalizing flows.
In other words, we will train multiple normalizing flows under identical conditions, just with different random initializations of the model parameters.
This will allow the optimization algorithm to explore different basins of attraction in the parameter space, allowing us to marginalize over different potential models.
In the machine learning literature, this is known as a Deep Ensemble \citep{lakshminarayanan2017}, a popular method for approximate bayesian marginalization \citep{wilson2020,fort2020}.

@@

For the training set, we use 100,000 galaxies from the catalog created in Section \ref{sec:galaxy-catalog}.
Each galaxy in the training set has a true redshift and observed magnitudes in the $ugrizy$ bands, with corresponding photometric errors.
To account for the photometric error, at the start of each training epoch, we resample the training set from the photometric error distribution.
In other words, each epoch, we resample
\begin{align}
    m \sim p(\hat{m}, \sigma_m),
\end{align}
where $\hat{m}$ are the observed magnitudes with photometric errors $\sigma_m$, and $p(\hat{m}, \sigma_m)$ is a Gaussian in flux space.
This allows our ensemble of flows to approximate the distribution $p(z, m)$.

@@

\begin{figure*}[t!]
    \script{photo-z/plot_ensemble_posteriors.py}
    \begin{centering}
        \includegraphics{figures/ensemble_posteriors.pdf}
        \caption{
            The ensemble of posteriors for three example galaxies.
            Flows 1-4 label the individual posteriors produced by each of the flows that make up the ensemble.
            The dashed black line is the mean of these individual posteriors and is the value used by the ensemble.
            The vertical gray line labeled ``Truth'' denotes the true redshift of the galaxy.
            Note the posteriors for most galaxies are not this broad or multimodal. 
        }
        \label{fig:ensemble-posteriors}
    \end{centering}
\end{figure*}

Each flow in the ensemble estimates the posterior for each galaxy by marginalizing over the photometric errors:
\begin{align}
    p(z| \hat{m}, \sigma_m) \propto \int dm \, p(z, m) \, p(m| \hat{m}, \sigma_m),
\end{align}
which is estimated by sampling $m \sim p(\hat{m}, \sigma_m)$ and averaging $p(z, m)$ over these samples.
We then average the posteriors from each flow in the ensemble.

An example of this for three galaxies can be seen in Figure \ref{fig:ensemble-posteriors}.
Each flow produces a PDF which may contain slightly different features in each case.
By averaging over the individual posteriors, we select for features that are common between models, while smoothing over features that are present in only a single model.
Compare these to the true posteriors.

We can calculate posteriors for galaxies with missing magnitudes by marginalizing over the missing magnitudes.
This is done by averaging $p(z, m)$ over a grid of values for the missing bands. 
This might be useful when not every galaxy is observed in every band, in which case you can marginalize over the bands that were not observed.
Or if a galaxy was observed but not detected in the band, you can marginalize over values for the band that are beyond the limiting magnitude.
You can also marginalize over bands for which you have a magnitude to understand what information about redshift is included in that band.

\note{Think about what I want to do here.}
For example, we randomly select \note{N} galaxies and estimate redshift posteriors while marginalizing over the $u$ band.
These posteriors can be seen in Figure \note{N}.
Mention how the loss of information degrades the posteriors.

@@

\begin{figure}[t!]
    \script{photo-z/plot_pz_point_estimates.py}
    \begin{centering}
        \includegraphics{figures/pz_point_estimates.pdf}
        \caption{
            Comparing the photo-z point estimates to the true redshifts in the catalog.
            The black diagonal line represents perfect photo-z performance.
        }
        \label{fig:point-estimates}
    \end{centering}
\end{figure}

The most common metrics for photo-z estimation concern photo-z point estimates, which are a compression of the photo-z posterior to a single redshift estimate.
We make the common choice of selecting the mode of the posteriors\footnote{The mean redshift is often a poor choice, since photo-z posteriors are generally multimodal, and so the mean value can lie between two modes at a redshift with very small probability density.}.
We compute metrics of the quantity $\Delta z = (z_\text{phot} - z_\text{true}) / (1 + z_\text{true})$, where the denominator accounts for naturally greater uncertainties at high redshift.

\begin{figure*}[t!]
    \script{photo-z/plot_binned_metrics.py}
    \begin{centering}
        \includegraphics{figures/binned_metrics.pdf}
        \caption{
            The bias, scatter, and catastrophic outlier fraction of the photo-z point estimates as a function of true galaxy redshift.
            The dashed black lines represent the science requirements for LSST cosmology as stated in the LSST SRD \needscite.
            These lines are to provide a sense of scale for these metrics.
            You can see that PZFlow meets the bias and scatter requirements up to redshift $\sim$ 1.5, while meeting the outlier fraction requirements for all redshifts.
            We note that individual redshifts do not actually need to meet the bias requirement as long as the bias can be well calibrated via some other source, e.g. galaxy clustering.
        }
        \label{fig:binned-metrics}
    \end{centering}
\end{figure*}

Figure \ref{fig:binned-metrics} shows the metrics from \citet{descSRD} as a function of true redshift.
The bias is defined as the median of $\Delta z$; the scatter is defined as $\text{IQR} / 1.349$, where IQR is the interquartile range of $\Delta z$; the outlier fraction is defined as the fraction of galaxies for which $\Delta z$ is greater than three times the $\Delta z$ scatter.
As a point of comparison, we plot the requirements for LSST cosmology in gray.

Like many photo-z estimators, our estimator performs well to a redshift of approximately 1.5.
At higher redshifts, our estimator does not meet the bias and scatter requirements, because there is very little training data in this redshift range.
We note however that for many cosmology application, it is okay for the bias to exceed the required limits, as long as the bias can be well determined via some calibration process \citep{newman2022}.

\begin{figure}[t!]
    \script{photo-z/plot_pit_histogram.py}
    \begin{centering}
        \includegraphics{figures/pit_histogram.pdf}
        \caption{
            The histogram of the probability integral transform (PIT) values of the photo-z posteriors.
            This figure characterizes the calibration of the estimated posteriors.
            In other words, does the true redshift lie at the center/in the tails of the posteriors as often as one would expect?
            Ideally, this histogram should be flat, matching the dashed black line.
            The fact that this distribution has a central peak indicates that the flow ensemble produces posteriors that are a little too conservative in their predictions.
            The slight rightward tilt indicates that the posteriors also have a negative redshift bias.
        }
        \label{fig:pit-histogram}
    \end{centering}
\end{figure}

Another common metric is the probability integral transform (PIT) (see e.g. \citealt{schmidt2020, dey2022}).
The PIT metric is used to determine if posteriors are well calibrated, i.e. if true values fall within X\% confidence intervals X\% of the time.
The PIT histogram for our estimator is shown in Figure \ref{fig:pit-histogram}.
Ideally, this histogram would be uniform and match the dashed horizontal line.
The fact that the histogram bulges at the center indicates that our estimator is too conservative -- i.e. the posteriors it produces are too broad.
This can be explained by the fact that normalizing flows exhibit mode covering behavior (the opposite of the mode collapse seen in GANsl; \citealt{salimans2016}).
In other words, because normalizing flows are trained via maximum likelihood estimation on the training samples, they receive very high penalties for missing any modes in the data.
As a result, they tend to conservatively spread out their density, in order to avoid any missed modes.
This results in overly conservative posterior predictions.
The low values at the edges of the PIT histogram indicate the relative rarity of catastrophic outliers, which is also reflected in the far right panel of Figure \ref{fig:binned-metrics}, where you can see that our estimator meets the requirement on the outlier fraction at all redshifts.

The previous metrics analyze photo-z performance for point estimates, which are insufficient for modern cosmology \citep{newman2022}, and for ensembles of posteriors, which is often not a good indicator of photo-z performance for science applications \citep{schmidt2020}.
One of the purposes of this paper is to create a galaxy catalog for which each galaxy has a true redshift posterior, which will enable more comprehensive evaluation of photo-z estimators.

\note{NEED TO FINISH WRITING THIS}
We will not undertake a full analysis of the posteriors here, but will compute a few metrics to point in that direction.
Calculate the KLD and maybe something else?
Point towards RAIL.

\section{Conclusion}

@@

\begin{figure*}[t!]
    \script{forward_model/plot_galaxy_losses.py}
    \begin{centering}
        \includegraphics{figures/galaxy_flow_losses.pdf}
        \caption{
            Training losses for the galaxy flows.
            After epoch 50 and 100, you can see a drop in the loss due to the decrease in the learning rate.
        }
        \label{fig:galaxy-flow-losses}
    \end{centering}
\end{figure*}

\note{the main flow:}
We trained the normalizing flow for 150 epochs using the Adam optimizer \citep{adam} (with a learning rate of $10^{-3}$, that decreased by a factor of 10 every 50 epochs) to minimize the negative log-likelihood,
\begin{align}
    \mathcal{L} = - \, \mathbb{E}[ \, \log p(x) \, ],
\end{align}
where the expectation is performed over galaxies in the training set and $p(x)$ is defined in Equation \ref{eq:px}.
Training took 7 minutes on a Tesla P100 12GB GPU.

\note{
    Need to add the curves for the ensembles.
    Move discussion of epochs and learning rates here too?    
}

@@

We estimate photometric errors for LSST using the error model from \citet{ivezic2019}, except we do not make the high SNR approximation used there.

\citet{ivezic2019} calculates a Gaussian photometric error in magnitude space using Equation 5 from that paper: 
\begin{align}
    \sigma_\text{rand}^2 = (0.04 - \gamma) x + \gamma x^2 ~~ (\text{mag}^2),
\end{align}
where $x \equiv 10 ^ {0.4(m - m_5)}$, $m_5$ is the 5$\sigma$ depth for the band, and $\gamma$ is a band-dependent parameter listed in Table 2 of the same paper.
This formula is based on the approximation
\begin{align}
    \sigma_\text{rand} = 2.5 \log_{10} \left(1 + \frac{N}{S} \right) \approx \frac{N}{S},
\end{align}
where $S$ and $N$ are the flux signal and noise, respectively.
This approximation is valid in the high signal-to-noise regime.

To generalize this model to the low signal-to-noise regime, we abandon this latter approximation.
For each magnitude, $m$, we calculate the corresponding flux, $S = 10^{-m/2.5}$, and the Gaussian flux noise,
\begin{align}
    N = S \, \sqrt{(0.04 - \gamma) x + \gamma x^2}, \label{eq:noise}.
\end{align}
For very bright galaxies, we impose a systematic floor corresponding to an error of 0.005 mags.
We then sample an observed flux, $S_\text{obs} \sim \mathcal{N}(S, N)$, and calculate the observed magnitude, $m_\text{obs} = -2.5 \log_{10} S_\text{obs}$.

For the reported photometric error, we do not use $N$, as this provides a deterministic link to the true magnitude via Equation \ref{eq:noise}.
Instead, we apply Equation \ref{eq:noise} again, this time using the observed magnitude instead of the true magnitude, yielding the estimated noise, $N_\text{obs}$.
Finally, we report the photometric error in magnitudes:
\begin{align}
    \sigma_m = 2.5 \log_{10} \left(1 + \frac{N_\text{obs}}{S_\text{obs}} \right).
\end{align}

\bibliography{bib.bib}
